---
title: "Gopher Tortoise Soil Biodiversity Study"
author: "Rachel Shepherd"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#About the analyses

These analyses are coupled with the manuscript "Gopher Tortoise Influence on Soil Biodiversity is Constrained by Historical Anthropogenic Land Use". 

##Load in packages 
```{r packages}
library(mctoolsr)
library(tidyverse)
library(emmeans)
library(lme4)
library(lmerTest)
library(effectsize)
library(corrplot)
library(vegan)
library(compositions)
```

##Load in data
Here we are using mctoolsr to manage the data
link to the source code: 
https://github.com/leffj/mctoolsr

The resulting files from the function load_taxa_table() are large lists. 
Each of these files is connected by matching sample IDs and taxonomy (ASV ID). 
Here is a breakdown of each file:  
data_loaded = ASV ID (row names) by sample ID (column names)  
map_loaded = sample ID (row names) by metadata attributes (column names)  
taxonomy_loaded = ASV ID (row names) by taxonomic strings (column names)  

```{r data}
# call .txt files
tax_table_nem = 'nem_asv_gt.txt'
tax_table_bact= 'bact_asv_gt.txt'
tax_table_fung= 'fung_asv_gt.txt'
tax_table_prot = 'prot_asv_gt_n.txt'
map_fp = 'metadata_gt_f.txt'

# load and filter for each broad group
input_nem = load_taxa_table(tax_table_nem, map_fp)
input_nem$map_loaded$sampleID <- row.names(input_nem$map_loaded)
input_nem_ra = convert_to_relative_abundances(input_nem)

input_bact = load_taxa_table(tax_table_bact, map_fp)
input_bact$map_loaded$sampleID <- row.names(input_bact$map_loaded)
input_bact_ra = convert_to_relative_abundances(input_bact)
input_bact_ra = filter_taxa_from_input(input_bact_ra, taxa_to_remove = "mitochondria")
input_bact_ra = filter_taxa_from_input(input_bact_ra, taxa_to_remove = "Chloroplast")

input_fung = load_taxa_table(tax_table_fung, map_fp)
input_fung$map_loaded$sampleID <- row.names(input_fung$map_loaded)
input_fung_ra = convert_to_relative_abundances(input_fung)


input_prot = load_taxa_table(tax_table_prot, map_fp)
input_prot$map_loaded$sampleID <- row.names(input_prot$map_loaded)
input_prot_ra = convert_to_relative_abundances(input_prot)

#plant data
plant_wide_n<-read.csv("plant_data_wide.csv", row.names = 1)
```

##First analyses: look at metadata (abiotic factors)
###Checking for correlations between factors
```{r correlation check}
# using the map loaded from the nematode large file
map = input_nem_ra$map_loaded

cormap<-map[,7:41]
M=cor(cormap,use="pairwise.complete.obs")
corrplot(M, method = 'number', type = 'upper',number.cex = 0.30,tl.cex = 0.75)
```
###After deciding which factors to keep, check normailty, run linear models, and make figures
```{r abiotic models and figures}
# check normality
hist(map$Soil.pH)
shapiro.test(map$Soil.pH)
hist(map$Organic.Matter)
shapiro.test(map$Organic.Matter) # not normal
hist(map$Total.Organic.C)
shapiro.test(map$Total.Organic.C) # not normal
hist(map$Nitrate)
shapiro.test(map$Nitrate) # not normal
hist(map$Ammonium)
shapiro.test(map$Ammonium)
hist(map$Total.Phosphorus)
shapiro.test(map$Total.Phosphorus) # not normal
hist(map$Percent.Moisture)
shapiro.test(map$Percent.Moisture) # not normal
hist(map$Calcium)
shapiro.test(map$Calcium)  # not normal

# go through all variables using lmer and anova 
biochem_models <- list(
  pH.lm = lmer(Soil.pH ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  log.OM.lm = lmer(log10(Organic.Matter) ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  log.oC.lm = lmer(log10(Total.Organic.C) ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  log.NO3.lm = lmer(log10(Nitrate) ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  NH4.lm = lmer(Ammonium ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  log.tP.lm = lmer(log10(Total.Phosphorus) ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  log.H2O.lm = lmer(log10(Percent.Moisture) ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  log.Ca.lm = lmer(log10(Calcium) ~ habitat*gopher_tortoise_activity +(1|site), data = map)
)

biochem_results <- do.call(rbind, lapply(names(biochem_models), function(resp) {

  aov_tab <- anova(biochem_models[[resp]])

  data.frame(
    response = resp,
    term     = rownames(aov_tab),
    NumDF    = aov_tab$NumDF,
    DenDF    = aov_tab$DenDF,
    F_value  = aov_tab$`F value`,
    p_value  = aov_tab$`Pr(>F)`
  )
}))

biochem_results$p_fdr <- NA

for (t in unique(biochem_results$term)) {
  idx <- biochem_results$term == t
  biochem_results$p_fdr[idx] <- p.adjust(biochem_results$p_value[idx], method = "fdr")
}

write.csv(biochem_results, "biochem_model_output.csv")

## make figures
#pH
map$treatment = factor(map$treatment, 
                    levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))


ggplot(map, aes(x = treatment, y = Soil.pH)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "pH",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

# Organic Matter
ggplot(map, aes(x = treatment, y = Organic.Matter)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Organic Matter",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

# Organic C
ggplot(map, aes(x = treatment, y = Total.Organic.C)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Organic Carbon",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

# NO3
ggplot(map, aes(x = treatment, y = Nitrate)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "NO3",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

# NH4
ggplot(map, aes(x = treatment, y = Ammonium)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "NH4",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

# Total P
ggplot(map, aes(x = treatment, y = Total.Phosphorus)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Total Phosphorus",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

# H2O
ggplot(map, aes(x = treatment, y = Percent.Moisture)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Soil Moisture",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

# Ca
ggplot(map, aes(x = treatment, y = Calcium)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Calcium",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))
```
##Analyses for the plant communities
###Alpha diversity
```{r plant alpha diversity}
# pull out mapping/metadata info
plant_mapping <- plant_wide_n[,1:4]

#rotate for diversity calculations
div_calcs <- as.data.frame(t(plant_wide_n))
colnames(div_calcs) <- div_calcs[1,]
div_calcs_f <- div_calcs[5:88,]
div_calcs_f[1:89] <- as.numeric(unlist(div_calcs_f[1:89]))

plant_div = calc_diversity(div_calcs_f, metric = "shannon")                   
plant_rich = calc_diversity(div_calcs_f, metric = "richness")

plant_mapping$plant_div = plant_div
plant_mapping$plant_rich = plant_rich

# check normality
hist(plant_mapping$plant_div)
shapiro.test(plant_mapping$plant_div) # not normal
hist(plant_mapping$plant_rich)
shapiro.test(plant_mapping$plant_rich) # not normal


# go through all variables using lmer and anova 
plant_models <- list(
  Plant.div.lm = lmer(plant_div ~ habitat + (1|site), data = plant_mapping),
  Plant.rich.lm = lmer(plant_rich ~ habitat + (1|site), data = plant_mapping)
  )

plant_results <- do.call(rbind, lapply(names(plant_models), function(resp) {

  aov_tab <- anova(plant_models[[resp]])

  data.frame(
    response = resp,
    term     = rownames(aov_tab),
    NumDF    = aov_tab$NumDF,
    DenDF    = aov_tab$DenDF,
    F_value  = aov_tab$`F value`,
    p_value  = aov_tab$`Pr(>F)`
  )
}))

plant_results$p_fdr <- NA

for (t in unique(plant_results$term)) {
  idx <- plant_results$term == t
  plant_results$p_fdr[idx] <- p.adjust(plant_results$p_value[idx], method = "fdr")
}

write.csv(plant_results, "plant_model_output.csv")

# plot
plant_mapping$habitat = factor(plant_mapping$habitat, 
                 levels = c("Intact", "Degraded"))

ggplot(plant_mapping, aes(x = habitat, y = plant_div)) +
  geom_boxplot(aes(fill = habitat), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = habitat),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("green4","green4")) +
  labs(x = NULL,
       y = "Shannon Diversity",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(plant_mapping, aes(x = habitat, y = plant_rich)) +
  geom_boxplot(aes(fill = habitat), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = habitat),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("green4","green4")) +
  labs(x = NULL,
       y = "Richness",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))
```
###Beta diversity
```{r plant beta diversity}
## plant species ndms
ord_plant <- metaMDS(t(div_calcs_f), distance = "bray", k = 3,
                   trymax = 200)
stressplot(ord_plant)
nmds1 <- ord_plant$points[,1]
nmds2 <- ord_plant$points[,2]
ord_plant_plot <- as.data.frame(cbind(nmds1,nmds2))


plant_mapping$site = factor(plant_mapping$site, 
                                     levels = c("1", "2", "3", "4",
                                                "5", "6", "7", "8", "9",
                                                "10", "11", "12", "13", "14", "15"))

# run permanova
plant_dist_bray <- vegdist(t(div_calcs_f), method = "euclidean", diag = FALSE, upper = TRUE,na.rm = TRUE)

with(plant_mapping, adonis2(plant_dist_bray ~ habitat, data = plant_mapping, permutations = 199, strata = site, by = "terms"))

# test homogeneity of dispersion among habitats
plant_disp <- betadisper(plant_dist_bray, plant_mapping$habitat)

# permutation test
anova(plant_disp)

# optional: pairwise test
permutest(plant_disp, pairwise = TRUE)

# plot
plant_mapping$habitat = factor(plant_mapping$habitat, 
                                     levels = c("Intact", "Degraded"))

ggplot(ord_plant_plot, aes(nmds1, nmds2, group = plant_mapping$habitat)) +
  geom_point(aes(shape = plant_mapping$habitat),
             size = 4, 
             stroke=0.25,
             fill = "green4") +
  scale_shape_manual(values = c(21,25)) +
  theme_classic() + 
  labs(x="NMDS 1", y="NMDS 2") 
```
##Plant functional group analysis (beta-diversity)
```{r plant functional group beta diversity}
# extract species coordinates
species_coords <- as.data.frame(scores(ord_plant, display = "species"))

# make row names
species_coords$species <- rownames(species_coords)

# merge plant species with functional traits 
trait_data <- read.csv("plant_functional_traits.csv")
species_with_traits <- left_join(species_coords, trait_data, by = "species")


ggplot() +
  geom_point(data = species_with_traits,
            aes(NMDS1, NMDS2, color = functional_group),
            size = 4, shape=8) +
  scale_color_manual(values = c("Forb" = "orange",
                                "Grass" = "maroon",
                                "Legume" = "lightblue3", 
                                "Woody" = "black")) +
  theme_classic() +
  labs(x = "NMDS 1", y = "NMDS 2")
```
##permanova for plant functional groups
```{r plant functional groups permanove}
plant_row_names <- plant_wide_n$plant_ID
plant_simplified = plant_wide_n[,5:80]
rownames(plant_simplified) = plant_row_names

meta <- data.frame(
  sample = rownames(plant_simplified),
  habitat = sub("\\..*$", "", rownames(plant_simplified)),   # F or S
  site = sub("^[A-Z]\\.(\\d+).*", "\\1", rownames(plant_simplified))
)

meta$site <- factor(meta$site)
meta$habitat <- factor(meta$habitat)
rownames(meta) <- meta$sample

fg_lookup <- trait_data

long_test <- plant_simplified %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  pivot_longer(
    -sample,
    names_to = "species",
    values_to = "abundance"
  ) %>%
  left_join(fg_lookup, by = "species")

fg_abund <- long_test %>%
  filter(!is.na(functional_group)) %>%          # drop unmatched species
  group_by(sample, functional_group) %>%
  summarise(
    abundance = sum(abundance),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = functional_group,
    values_from = abundance,
    values_fill = 0
  )

fg_mat <- fg_abund %>%
  column_to_rownames("sample") %>%
  as.matrix()

# run permanova
plant_func_dist_bray <- vegdist(fg_mat, method = "euclidean", diag = FALSE, upper = TRUE, na.rm = TRUE)
adonis2(plant_func_dist_bray ~ habitat, data = meta, permutations = 199, strata = meta$site, by = "terms")

```

##Qubit plot for microbial biomass
```{r qubit plot microbial biomass}
# plot
map$treatment = factor(map$treatment, 
                       levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))


ggplot(map, aes(x = treatment, y = Qubit)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Qubit",
       title = NULL) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))
```
##Analyses for soil alpha diversity metrics 
##Starting with microbial alpha diversity analyses
```{r microbial alhpa diversity models and figures}
# make new map file
map = input_nem_ra$map_loaded
# calculate diversity metrics and add to nematode mapping file
bact_rich = calc_diversity(input_bact_ra$data_loaded, metric = 'richness')
bact_div = calc_diversity(input_bact_ra$data_loaded, metric = 'shannon')
map$bact_rich = bact_rich
map$bact_div = bact_div
fung_rich = calc_diversity(input_fung_ra$data_loaded, metric = 'richness')
fung_div = calc_diversity(input_fung_ra$data_loaded, metric = 'shannon')
map$fung_rich = fung_rich
map$fung_div = fung_div
prot_rich = calc_diversity(input_prot_ra$data_loaded, metric = 'richness')
prot_div = calc_diversity(input_prot_ra$data_loaded, metric = 'shannon')
map$prot_rich = prot_rich 
map$prot_div = prot_div

# exclude outlier protist sample from analyses
protist_set_for_alpha_div = map %>%
  filter(sampleID != "S.9.A_F")

# models for microbial biomass (quibit) richness and diversity 
# go through all variables using lmer and anova 
microbial_models <- list(
  Qubit.lm = lmer(Qubit ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  bact.rich.lm = lmer(bact_rich ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  bact.div.lm = lmer(bact_div ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  fung.rich.lm = lmer(fung_rich ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  fung.div.lm = lmer(fung_div ~ habitat*gopher_tortoise_activity +(1|site), data = map),
  prot.rich.lm = lmer(prot_rich ~ habitat*gopher_tortoise_activity +(1|site), data = protist_set_for_alpha_div),
  prot.div.lm = lmer(prot_div ~ habitat*gopher_tortoise_activity +(1|site), data = protist_set_for_alpha_div)
  )

microbial_results <- do.call(rbind, lapply(names(microbial_models), function(resp) {

  aov_tab <- anova(microbial_models[[resp]])

  data.frame(
    response = resp,
    term     = rownames(aov_tab),
    NumDF    = aov_tab$NumDF,
    DenDF    = aov_tab$DenDF,
    F_value  = aov_tab$`F value`,
    p_value  = aov_tab$`Pr(>F)`
  )
}))

microbial_results$p_fdr <- NA

for (t in unique(microbial_results$term)) {
  idx <- microbial_results$term == t
  microbial_results$p_fdr[idx] <- p.adjust(microbial_results$p_value[idx], method = "fdr")
}

write.csv(microbial_results, "microbial_model_output.csv")

# plot
map$treatment = factor(map$treatment, 
                 levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))

#plot bacterial richness
ggplot(map, aes(x = treatment, y = bact_rich)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Richness",
       title = "Bacteria") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))


#plot bacterial diversity
ggplot(map, aes(x = treatment, y = bact_div)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Shannon Diversity",
       title = "Bacteria") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

#plot fungal richness
ggplot(map, aes(x = treatment, y = fung_rich)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Richness",
       title = "Fungi") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))


#plot fungal diversity
ggplot(map, aes(x = treatment, y = fung_div)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Shannon Diversity",
       title = "Fungi") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

#plot protist richness
map %>% 
  filter(sampleID != "S.9.A_F") %>%
ggplot(aes(x = treatment, y = prot_rich)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Richness",
       title = "Protist") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))


#plot protist diversity
map %>% 
  filter(sampleID != "S.9.A_F") %>%
ggplot(aes(x = treatment, y = prot_div)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Shannon Diversity",
       title = "Protist") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))
```
##Look at the effect size of gopher tortoise activity on diversity
```{r microbial alpha diversity effect sizes}
# bacteria
bact_map_intact = map %>%
  filter(habitat == "Intact")
bact_map_intact$gopher_tortoise_activity = factor(bact_map_intact$gopher_tortoise_activity, 
                                levels = c("Foraging", "Burrowing"))

cohens_d(bact_rich ~ gopher_tortoise_activity, data = bact_map_intact)
cohens_d(bact_div ~ gopher_tortoise_activity, data = bact_map_intact)

bact_map_degraded = map %>%
  filter(habitat == "Degraded")
bact_map_degraded$gopher_tortoise_activity = factor(bact_map_degraded$gopher_tortoise_activity, 
                                  levels = c("Foraging", "Burrowing"))

cohens_d(bact_rich ~ gopher_tortoise_activity, data = bact_map_degraded)
cohens_d(bact_div ~ gopher_tortoise_activity, data = bact_map_degraded)

# fungi
fung_map_intact = map %>%
  filter(habitat == "Intact")
fung_map_intact$gopher_tortoise_activity = factor(fung_map_intact$gopher_tortoise_activity, 
                                levels = c("Foraging", "Burrowing"))

cohens_d(fung_rich ~ gopher_tortoise_activity, data = fung_map_intact)
cohens_d(fung_div ~ gopher_tortoise_activity, data = fung_map_intact)

fung_map_degraded = map %>%
  filter(habitat == "Degraded")
fung_map_degraded$gopher_tortoise_activity = factor(fung_map_degraded$gopher_tortoise_activity, 
                                  levels = c("Foraging", "Burrowing"))

cohens_d(fung_rich ~ gopher_tortoise_activity, data = fung_map_degraded)
cohens_d(fung_div ~ gopher_tortoise_activity, data = fung_map_degraded)

# protist
prot_map_intact = map %>%
  filter(habitat == "Intact")
prot_map_intact$gopher_tortoise_activity = factor(prot_map_intact$gopher_tortoise_activity, 
                                levels = c("Foraging", "Burrowing"))

cohens_d(prot_rich ~ gopher_tortoise_activity, data = prot_map_intact)
cohens_d(prot_div ~ gopher_tortoise_activity, data = prot_map_intact)

prot_map_degraded = map %>%
  filter(habitat == "Degraded")
prot_map_degraded$gopher_tortoise_activity = factor(prot_map_degraded$gopher_tortoise_activity, 
                                  levels = c("Foraging", "Burrowing"))

cohens_d(prot_rich ~ gopher_tortoise_activity, data = prot_map_degraded)
cohens_d(prot_div ~ gopher_tortoise_activity, data = prot_map_degraded)
```
##Alpha diversity for nematode trophic groups
###run models and make figures
```{r nematode alpha diversity models and figures}
# Bacterial-Feeding
BF_alpha = filter_taxa_from_input(input_nem_ra, at_spec_level = 3,
                                  taxa_to_keep = "BF")
BF_rich = calc_diversity(BF_alpha$data_loaded, metric = 'richness')
BF_div = calc_diversity(BF_alpha$data_loaded, metric = 'shannon')

# join to map
BF_map = BF_alpha$map_loaded
BF_map$BF_rich = BF_rich
BF_map$BF_div = BF_div

# Fungal-feeding
FF_alpha = filter_taxa_from_input(input_nem_ra, at_spec_level = 3,
                                  taxa_to_keep = "FF")
FF_rich = calc_diversity(FF_alpha$data_loaded, metric = 'richness')
FF_div = calc_diversity(FF_alpha$data_loaded, metric = 'shannon')

# join to map
FF_map = FF_alpha$map_loaded
FF_map$FF_rich = FF_rich
FF_map$FF_div = FF_div

# Plant-parasitic
PP_alpha = filter_taxa_from_input(input_nem_ra, at_spec_level = 3,
                                  taxa_to_keep = "PP")
PP_rich = calc_diversity(PP_alpha$data_loaded, metric = 'richness')
PP_div = calc_diversity(PP_alpha$data_loaded, metric = 'shannon')

# join to map
PP_map = PP_alpha$map_loaded
PP_map$PP_rich = PP_rich
PP_map$PP_div = PP_div

# Omnivorous
OM_alpha = filter_taxa_from_input(input_nem_ra, at_spec_level = 3,
                                  taxa_to_keep = "OM")
OM_rich = calc_diversity(OM_alpha$data_loaded, metric = 'richness')
OM_div = calc_diversity(OM_alpha$data_loaded, metric = 'shannon')

# join to map
OM_map = OM_alpha$map_loaded
OM_map$OM_rich = OM_rich
OM_map$OM_div = OM_div

# Root-associated
RA_alpha = filter_taxa_from_input(input_nem_ra, at_spec_level = 3,
                                  taxa_to_keep = "RA")
RA_rich = calc_diversity(RA_alpha$data_loaded, metric = 'richness')
RA_div = calc_diversity(RA_alpha$data_loaded, metric = 'shannon')

# join to map
RA_map = RA_alpha$map_loaded
RA_map$RA_rich = RA_rich
RA_map$RA_div = RA_div

# models for nematode richness and diversity 
# go through all variables using lmer and anova 
nematode_models <- list(
  BF.rich.lm = lmer(BF_rich ~ gopher_tortoise_activity * habitat +(1|site), data = BF_map),
  BF.div.lm = lmer(BF_div ~ gopher_tortoise_activity * habitat +(1|site), data = BF_map),
  FF.rich.lm = lmer(FF_rich ~ gopher_tortoise_activity * habitat +(1|site), data = FF_map),
  FF.div.lm = lmer(FF_div ~ gopher_tortoise_activity * habitat +(1|site), data = FF_map),
  PP.rich.lm = lmer(PP_rich ~ gopher_tortoise_activity * habitat +(1|site), data = PP_map),
  PP.div.lm = lmer(PP_div ~ gopher_tortoise_activity * habitat +(1|site), data = PP_map),
  OM.rich.lm = lmer(OM_rich ~ gopher_tortoise_activity * habitat +(1|site), data = OM_map),
  OM.div.lm = lmer(OM_div ~ gopher_tortoise_activity * habitat +(1|site), data = OM_map),
  RA.rich.lm = lmer(RA_rich ~ gopher_tortoise_activity * habitat +(1|site), data = RA_map),
  RA.div.lm = lmer(RA_div ~ gopher_tortoise_activity * habitat +(1|site), data = RA_map)
  )

nematode_results <- do.call(rbind, lapply(names(nematode_models), function(resp) {

  aov_tab <- anova(nematode_models[[resp]])

  data.frame(
    response = resp,
    term     = rownames(aov_tab),
    NumDF    = aov_tab$NumDF,
    DenDF    = aov_tab$DenDF,
    F_value  = aov_tab$`F value`,
    p_value  = aov_tab$`Pr(>F)`
  )
}))

nematode_results$p_fdr <- NA

for (t in unique(nematode_results$term)) {
  idx <- nematode_results$term == t
  nematode_results$p_fdr[idx] <- p.adjust(nematode_results$p_value[idx], method = "fdr")
}

write.csv(nematode_results, "nematode_div_model_output.csv")


# Plot all trophic groups
BF_map$treatment = factor(BF_map$treatment, 
                     levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))


ggplot(BF_map, aes(x = treatment, y = BF_rich)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Richness",
       title = "Bacterial-Feeding") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(BF_map, aes(x = treatment, y = BF_div)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Shannon Diversity",
       title = "Bacterial-Feeding") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

FF_map$treatment = factor(FF_map$treatment, 
                    levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))


ggplot(FF_map, aes(x = treatment, y = FF_rich)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Richness",
       title = "Fungal-Feeding") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(FF_map, aes(x = treatment, y = FF_div)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Shannon Diversity",
       title = "Fungal-Feeding") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

PP_map$treatment = factor(PP_map$treatment, 
                    levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))


ggplot(PP_map, aes(x = treatment, y = PP_rich)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Richness",
       title = "Plant-Parasitic") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(PP_map, aes(x = treatment, y = PP_div)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Shannon Diversity",
       title = "Plant-Parasitic") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

# plot
OM_map$treatment = factor(OM_map$treatment, 
                    levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))


ggplot(OM_map, aes(x = treatment, y = OM_rich)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Richness",
       title = "Omnivorous") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(OM_map, aes(x = treatment, y = OM_div)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Shannon Diversity",
       title = "Omnivorous") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

# plot
RA_map$treatment = factor(RA_map$treatment, 
                    levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))


ggplot(RA_map, aes(x = treatment, y = RA_rich)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Richness",
       title = "Root-Associated") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))



ggplot(RA_map, aes(x = treatment, y = RA_div)) +
  geom_boxplot(aes(fill = treatment), outlier.shape = NA,
               color = c("black")) +    
  geom_jitter(aes(fill = treatment),
              alpha = 0.8,
              shape = 21,
              size = 3,
              set.seed(666),
              color = "black") +
  scale_fill_manual(values = c("khaki", "green4","khaki","green4")) +
  labs(x = NULL,
       y = "Shannon Diversity",
       title = "Root-Associated") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5))

```
##Nematode alpha diversity effect sizes
```{r nematode effect sizes}
# effect size
BF_map_intact = BF_map %>%
  filter(habitat == "Intact")
BF_map_intact$gopher_tortoise_activity = factor(BF_map_intact$gopher_tortoise_activity, 
                           levels = c("Burrowing", "Foraging"))

cohens_d(BF_rich ~ gopher_tortoise_activity, data = BF_map_intact)
cohens_d(BF_div ~ gopher_tortoise_activity, data = BF_map_intact)

BF_map_degraded = BF_map %>%
  filter(habitat == "Degraded")
BF_map_degraded$gopher_tortoise_activity = factor(BF_map_degraded$gopher_tortoise_activity, 
                               levels = c("Burrowing", "Foraging"))

cohens_d(BF_rich ~ gopher_tortoise_activity, data = BF_map_degraded)
cohens_d(BF_div ~ gopher_tortoise_activity, data = BF_map_degraded)

# effect size
FF_map_intact = FF_map %>%
  filter(habitat == "Intact")
FF_map_intact$gopher_tortoise_activity = factor(FF_map_intact$gopher_tortoise_activity, 
                               levels = c("Burrowing", "Foraging"))

cohens_d(FF_rich ~ gopher_tortoise_activity, data = FF_map_intact)
cohens_d(FF_div ~ gopher_tortoise_activity, data = FF_map_intact)

FF_map_degraded = FF_map %>%
  filter(habitat == "Degraded")
FF_map_degraded$gopher_tortoise_activity = factor(FF_map_degraded$gopher_tortoise_activity, 
                                 levels = c("Burrowing", "Foraging"))

cohens_d(FF_rich ~ gopher_tortoise_activity, data = FF_map_degraded)
cohens_d(FF_div ~ gopher_tortoise_activity, data = FF_map_degraded)

# effect size
PP_map_intact = PP_map %>%
  filter(habitat == "Intact")
PP_map_intact$gopher_tortoise_activity = factor(PP_map_intact$gopher_tortoise_activity, 
                               levels = c("Burrowing", "Foraging"))

cohens_d(PP_rich ~ gopher_tortoise_activity, data = PP_map_intact)
cohens_d(PP_div ~ gopher_tortoise_activity, data = PP_map_intact)

PP_map_degraded = PP_map %>%
  filter(habitat == "Degraded")
PP_map_degraded$gopher_tortoise_activity = factor(PP_map_degraded$gopher_tortoise_activity, 
                                 levels = c("Burrowing", "Foraging"))

cohens_d(PP_rich ~ gopher_tortoise_activity, data = PP_map_degraded)
cohens_d(PP_div ~ gopher_tortoise_activity, data = PP_map_degraded)

# effect size
OM_map_intact = OM_map %>%
  filter(habitat == "Intact")
OM_map_intact$gopher_tortoise_activity = factor(OM_map_intact$gopher_tortoise_activity, 
                               levels = c("Burrowing", "Foraging"))

cohens_d(OM_rich ~ gopher_tortoise_activity, data = OM_map_intact)
cohens_d(OM_div ~ gopher_tortoise_activity, data = OM_map_intact)

OM_map_degraded = OM_map %>%
  filter(habitat == "Degraded")
OM_map_degraded$gopher_tortoise_activity = factor(OM_map_degraded$gopher_tortoise_activity, 
                                 levels = c("Burrowing", "Foraging"))

cohens_d(OM_rich ~ gopher_tortoise_activity, data = OM_map_degraded)
cohens_d(OM_div ~ gopher_tortoise_activity, data = OM_map_degraded)

# effect size
RA_map_intact = RA_map %>%
  filter(habitat == "Intact")
RA_map_intact$gopher_tortoise_activity = factor(RA_map_intact$gopher_tortoise_activity, 
                               levels = c("Burrowing", "Foraging"))

cohens_d(RA_rich ~ gopher_tortoise_activity, data = RA_map_intact)
cohens_d(RA_div ~ gopher_tortoise_activity, data = RA_map_intact)

RA_map_degraded = RA_map %>%
  filter(habitat == "Degraded")
RA_map_degraded$gopher_tortoise_activity = factor(RA_map_degraded$gopher_tortoise_activity, 
                                 levels = c("Burrowing", "Foraging"))

cohens_d(RA_rich ~ gopher_tortoise_activity, data = RA_map_degraded)
cohens_d(RA_div ~ gopher_tortoise_activity, data = RA_map_degraded)
```
##Add mean plant diversity per site and microbial/nematode diversity to overall mapping file for downstream analyses
```{r adding diversities to mapping files}
# add new plant metrics to overall mapping file for 
plant_rich_site = plant_mapping %>%
  group_by(habitat, site) %>%
  summarise(mean_plant_rich = mean(plant_rich)) 

plant_div_site = plant_mapping %>%
  group_by(habitat, site) %>%
  summarise(mean_plant_div = mean(plant_div)) 

plant_distinct_plot = merge(plant_div_site, plant_rich_site)
shorten_eco <- c("Intact" = "S", "Degraded" = "F")
plant_distinct_plot$sampleID <- paste0(shorten_eco[plant_distinct_plot$habitat], ".", plant_distinct_plot$site, ".O_F")
plant_distinct_plot <- plant_distinct_plot[3:5]
map_w_div <- merge(map, plant_distinct_plot, 
                   by = "sampleID", all.x = TRUE)

# replace NAs (for burrowing plots) with 0
map_w_div$mean_plant_div[is.na(map_w_div$mean_plant_div)] <- 0
map_w_div$mean_plant_rich[is.na(map_w_div$mean_plant_rich)] <- 0


write.table(map_w_div, "metadata_gt_all_with_div.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```
##Reload all data with diversity metrics included
```{r reload all data with plant metrics}
# remove existing files
rm(list = ls())

# call .txt files
tax_table_nem = 'nem_asv_gt.txt'
tax_table_bact= 'bact_asv_gt.txt'
tax_table_fung= 'fung_asv_gt.txt'
tax_table_prot = 'prot_asv_gt_n.txt'
map_fp_plants = 'metadata_gt_all_with_div.txt'

# load and filter for each broad group
input_nem = load_taxa_table(tax_table_nem, map_fp_plants)
input_nem$map_loaded$sampleID <- row.names(input_nem$map_loaded)
input_nem_ra = convert_to_relative_abundances(input_nem)

input_bact = load_taxa_table(tax_table_bact, map_fp_plants)
input_bact$map_loaded$sampleID <- row.names(input_bact$map_loaded)
input_bact_ra = convert_to_relative_abundances(input_bact)
input_bact_ra = filter_taxa_from_input(input_bact_ra, taxa_to_remove = "mitochondria")
input_bact_ra = filter_taxa_from_input(input_bact_ra, taxa_to_remove = "Chloroplast")

input_fung = load_taxa_table(tax_table_fung, map_fp_plants)
input_fung$map_loaded$sampleID <- row.names(input_fung$map_loaded)
input_fung_ra = convert_to_relative_abundances(input_fung)


input_prot = load_taxa_table(tax_table_prot, map_fp_plants)
input_prot$map_loaded$sampleID <- row.names(input_prot$map_loaded)
input_prot_ra = convert_to_relative_abundances(input_prot)
```
##Run analyses on lineages (at varying taxonomic resolution) for each group
###Starting with bacteria
```{r bacterial relative abundances analyses}
# Summarize bacteria at the phylum level #
sumtax_bact_phylum <- summarize_taxonomy(input_bact_ra, level = 2, report_higher_tax = F)

# clean up names... create separate phylum level taxonomy
bact_phylum_taxonomy <- as.data.frame(rownames(sumtax_bact_phylum))
bact_phylum_taxonomy$label <- bact_phylum_taxonomy$`rownames(sumtax_bact_phylum)`

bact_phylum_taxonomy$label = recode(bact_phylum_taxonomy$label,
                               "Deinococcus-Thermus" = "Deinococcus_Thermus",
                               "BHI80-139" = "BHI80_139",
                               "NPL-UPA2" = "NPL_UPA2",
                               "WCHB1-60" = "WCHB1_60")
# go by labels, flip
rownames(sumtax_bact_phylum) <- bact_phylum_taxonomy$label
sumtax_bact_phylum.t <- as.data.frame(t(sumtax_bact_phylum))


#map
bact_map_model <- input_bact_ra$map_loaded

# merge phylum table and map
bact_phylum_wmap <- merge(sumtax_bact_phylum.t, bact_map_model, by = 0)

# convert to matrix for CLR 
bact_matrix <- as.matrix(bact_phylum_wmap[, 2:35])

# add small pseudocount
bact_clr <- clr(bact_matrix + 1e-6)

full_bact_clr <- bind_cols(
  bact_phylum_wmap[, -c(2:35)],
  as.data.frame(bact_clr)
)

# models for bacterial abundance
# go through all variables using lmer and anova 
bacteria_models <- list(
 Acidobacteria=lmer(Acidobacteria~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Actinobacteria=lmer(Actinobacteria~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Armatimonadetes=lmer(Armatimonadetes~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Bacteroidetes=lmer(Bacteroidetes~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
BHI80_139=lmer(BHI80_139~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Candidate_division_BRC1=lmer(Candidate_division_BRC1~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Candidate_division_OD1=lmer(Candidate_division_OD1~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Candidate_division_OP11=lmer(Candidate_division_OP11~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Candidate_division_OP3=lmer(Candidate_division_OP3~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Candidate_division_TM7=lmer(Candidate_division_TM7~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Candidate_division_WS3=lmer(Candidate_division_WS3~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Chlamydiae=lmer(Chlamydiae~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Chlorobi=lmer(Chlorobi~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Chloroflexi=lmer(Chloroflexi~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Cyanobacteria=lmer(Cyanobacteria~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Deinococcus_Thermus=lmer(Deinococcus_Thermus~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Elusimicrobia=lmer(Elusimicrobia~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Fibrobacteres=lmer(Fibrobacteres~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Firmicutes=lmer(Firmicutes~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
GAL08=lmer(GAL08~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Gemmatimonadetes=lmer(Gemmatimonadetes~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Lentisphaerae=lmer(Lentisphaerae~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Nitrospirae=lmer(Nitrospirae~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
NPL_UPA2=lmer(NPL_UPA2~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Planctomycetes=lmer(Planctomycetes~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Proteobacteria=lmer(Proteobacteria~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
SM2F11=lmer(SM2F11~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Spirochaetes=lmer(Spirochaetes~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
TA06=lmer(TA06~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Tenericutes=lmer(Tenericutes~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Thermotogae=lmer(Thermotogae~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
TM6=lmer(TM6~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
Verrucomicrobia=lmer(Verrucomicrobia~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr),
WCHB1_60=lmer(WCHB1_60~ habitat*gopher_tortoise_activity + (1|site), data = full_bact_clr)
  )

bacteria_results <- do.call(rbind, lapply(names(bacteria_models), function(resp) {

  aov_tab <- anova(bacteria_models[[resp]])

  data.frame(
    response = resp,
    term     = rownames(aov_tab),
    NumDF    = aov_tab$NumDF,
    DenDF    = aov_tab$DenDF,
    F_value  = aov_tab$`F value`,
    p_value  = aov_tab$`Pr(>F)`
  )
}))

bacteria_results$p_fdr <- NA

for (t in unique(bacteria_results$term)) {
  idx <- bacteria_results$term == t
  bacteria_results$p_fdr[idx] <- p.adjust(bacteria_results$p_value[idx], method = "fdr")
}

bacteria_results2 <- bacteria_results %>%
  mutate(term = recode(term,
    "gopher_tortoise_activity" = "Activity",
    "habitat" = "Vegetation Type",
    "habitat:gopher_tortoise_activity" = "Interaction"
  ))

write.csv(bacteria_results2, "raw_bact_model_output.csv")

bacteria_results2 <- bacteria_results2 %>%
  mutate(F_p = sprintf("F = %.2f, p = %.3f", F_value, p_fdr))

bacteria_table <- bacteria_results2 %>%
  select(response, term, F_p) %>%
  pivot_wider(
    names_from = term,
    values_from = F_p
  )

bacteria_table <- bacteria_table %>%
  rename(group = response)

write.csv(bacteria_table, "bacteria_model_output_clr.csv")
```

```{r bacteria relative abundance figure}
bact_map = input_bact_ra$map_loaded


## BY Treatment
table_bact_level2_fp_sort = summarize_taxonomy(input_bact_ra, level = 2,
                                              report_higher_tax = F)



total_bact_sort = table_bact_level2_fp_sort %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  left_join(bact_map) %>%
  pivot_longer(cols=c(2:35), names_to = "Phylum", values_to = 'abund') %>%
  as.data.frame() 

# make mean relative abundacne by treatment
trt_bact = total_bact_sort %>%
  group_by(treatment, Phylum) %>% 
  summarise(mean_RA = mean(abund))

trt_bact$Phylum = factor(trt_bact$Phylum, 
                        levels = c("Tenericutes",
                                   "Candidate_division_OP11",
                                   "Candidate_division_OD1",
                                   "Candidate_division_TM7",
                                   "Lentisphaerae",
                                   "Candidate_division_BRC1",
                                   "BHI80-139",
                                   "TA06",
                                   "Spirochaetes",
                                   "NPL-UPA2",
                                   "WCHB1-60",
                                   "SM2F11",
                                   "GAL08",
                                   "Fibrobacteres",
                                   "Deinococcus-Thermus",
                                   "TM6",
                                   "Chlamydiae",
                                   "Candidate_division_OP3",
                                   "Candidate_division_WS3",
                                   "Chlorobi",
                                   "Thermotogae",
                                   "Elusimicrobia",
                                   "Armatimonadetes",
                                   "Nitrospirae",
                                   "Gemmatimonadetes",
                                   "Bacteroidetes",
                                   "Cyanobacteria",
                                   "Chloroflexi",
                                   "Verrucomicrobia",
                                   "Planctomycetes",
                                   "Actinobacteria",
                                   "Firmicutes",
                                   "Acidobacteria",
                                   "Proteobacteria" ))

trt_bact$treatment = factor(trt_bact$treatment, 
                     levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))

ggplot(trt_bact, aes(x=treatment, y=mean_RA)) + 
  geom_bar(aes(fill = Phylum), position = "stack", stat="identity") +
  scale_fill_manual(values = c("#E0B0FF",	"#7F00FF",		
                               "#40E0D0",
                               "#00004d",	"#00FFFF",	"#000066",	"#0000FF",	"#0096FF",	"#6699ff",  "#0033ff",  "#3366ff",  "#1434A4",	
                               "#4d0000", "#b30000",  "#660000",  "#ff3333",  "#cc0000",  "#ff8080",  "#990000", 
                               "#BF40BF",	"#AA336A",  "#FF00FF",	"#FF69B4",	"#FFB6C1",  "#E30B5C",
                               "#FFBF00",	"#FBCEB1",	"#CD7F32",	"#FFC000",	"#FFEA00",	"#F28C28",	"#E97451",	"#FAD5A5",	"#DAA06D",	"#F4BB44",	"#FFAA33",	"#B87333",	"#FF7518",	"#CC5500",	"#FFFF8F",	"#F0E68C",	"#FADA5E",	"#FFFAA0",	"#F8DE7E",	"#FDDA0D",	"#E1C16E"))+
  labs(
    x = NULL,
    y = "Relative Abundance of Bacterial Phyla",
    fill = NULL
  ) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) 
```
###Fungi
```{r fungal relative abundances}
# Summarize fungi at the phylum level #
sumtax_fungi_phylum <- summarize_taxonomy(input_fung_ra, level = 7, report_higher_tax = F)

# clean up names... create separate phylum level taxonomy
fung_phylum_taxonomy <- as.data.frame(rownames(sumtax_fungi_phylum))
fung_phylum_taxonomy$label <- fung_phylum_taxonomy$`rownames(sumtax_fungi_phylum)`

# go by labels, flip
rownames(sumtax_fungi_phylum) <- fung_phylum_taxonomy$label
sumtax_fungi_phylum.t <- as.data.frame(t(sumtax_fungi_phylum))

#map
fung_map_model <- input_fung_ra$map_loaded

# merge phylum table and map
fung_phylum_wmap <- merge(sumtax_fungi_phylum.t, fung_map_model, by = 0)

# convert to matrix for CLR 
fung_matrix <- as.matrix(fung_phylum_wmap[, 2:9])

# add small pseudocount
fung_clr <- clr(fung_matrix + 1e-6)

full_fung_clr <- bind_cols(
  fung_phylum_wmap[, -c(2:9)],
  as.data.frame(fung_clr)
)

# models for fungal abundance
# go through all variables using lmer and anova 
fungal_models <- list(
Ascomycota=lmer(Ascomycota~ habitat*gopher_tortoise_activity + (1|site), data = full_fung_clr),
Basidiomycota=lmer(Basidiomycota~ habitat*gopher_tortoise_activity + (1|site), data = full_fung_clr),
Blastocladiomycota=lmer(Blastocladiomycota~ habitat*gopher_tortoise_activity + (1|site), data = full_fung_clr),
Chytridiomycota=lmer(Chytridiomycota~ habitat*gopher_tortoise_activity + (1|site), data = full_fung_clr),
Cryptomycota=lmer(Cryptomycota~ habitat*gopher_tortoise_activity + (1|site), data = full_fung_clr),
Mucoromycota=lmer(Mucoromycota~ habitat*gopher_tortoise_activity + (1|site), data = full_fung_clr),
Olpidiomycota=lmer(Olpidiomycota~ habitat*gopher_tortoise_activity + (1|site), data = full_fung_clr),
Zoopagomycota=lmer(Zoopagomycota~ habitat*gopher_tortoise_activity + (1|site), data = full_fung_clr)
 )

fungal_results <- do.call(rbind, lapply(names(fungal_models), function(resp) {

  aov_tab <- anova(fungal_models[[resp]])

  data.frame(
    response = resp,
    term     = rownames(aov_tab),
    NumDF    = aov_tab$NumDF,
    DenDF    = aov_tab$DenDF,
    F_value  = aov_tab$`F value`,
    p_value  = aov_tab$`Pr(>F)`
  )
}))

fungal_results$p_fdr <- NA

for (t in unique(fungal_results$term)) {
  idx <- fungal_results$term == t
  fungal_results$p_fdr[idx] <- p.adjust(fungal_results$p_value[idx], method = "fdr")
}

fungal_results2 <- fungal_results %>%
  mutate(term = recode(term,
    "gopher_tortoise_activity" = "Activity",
    "habitat" = "Vegetation Type",
    "habitat:gopher_tortoise_activity" = "Interaction"
  ))

write.csv(fungal_results2, "raw_fungal_model_output.csv")

fungal_results2 <- fungal_results2 %>%
  mutate(F_p = sprintf("F = %.2f, p = %.3f", F_value, p_fdr))

fungal_table <- fungal_results2 %>%
  select(response, term, F_p) %>%
  pivot_wider(
    names_from = term,
    values_from = F_p
  )

fungal_table <- fungal_table %>%
  rename(group = response)

write.csv(fungal_table, "fungal_model_output_clr.csv")
```

```{r fungal relative abundance figure}
fung_map = input_fung_ra$map_loaded

## BY Treatment
table_fung_level7_fp_sort = summarize_taxonomy(input_fung_ra, level = 7,
                                               report_higher_tax = F)



total_fung_sort = table_fung_level7_fp_sort %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  left_join(fung_map) %>%
  pivot_longer(cols=c(2:9), names_to = "Phylum", values_to = 'abund') %>%
  as.data.frame() 

# make mean RA by trt
trt_fung = total_fung_sort %>%
  group_by(treatment, Phylum) %>% 
  summarise(mean_RA = mean(abund))

trt_fung$Phylum = factor(trt_fung$Phylum, 
                         levels = c("Cryptomycota",
                                    "Blastocladiomycota",
                                    "Zoopagomycota",
                                    "Olpidiomycota",
                                    "Chytridiomycota",
                                    "Mucoromycota",
                                    "Ascomycota",
                                    "Basidiomycota"))

trt_fung$treatment = factor(trt_fung$treatment, 
                      levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))

ggplot(trt_fung, aes(x=treatment, y=mean_RA)) + 
  geom_bar(aes(fill = Phylum), position = "stack", stat="identity") +
  scale_fill_manual(values = c("#E0B0FF",	"#7F00FF",		
                               "#40E0D0",
                               "#00004d",	"#00FFFF",	"#000066",	"#0000FF",	"#0096FF",	"#6699ff",  "#0033ff",  "#3366ff",  "#1434A4",	
                               "#4d0000", "#b30000",  "#660000",  "#ff3333",  "#cc0000",  "#ff8080",  "#990000", 
                               "#BF40BF",	"#AA336A",  "#FF00FF",	"#FF69B4",	"#FFB6C1",  "#E30B5C",
                               "#FFBF00",	"#FBCEB1",	"#CD7F32",	"#FFC000",	"#FFEA00",	"#F28C28",	"#E97451",	"#FAD5A5",	"#DAA06D",	"#F4BB44",	"#FFAA33",	"#B87333",	"#FF7518",	"#CC5500",	"#FFFF8F",	"#F0E68C",	"#FADA5E",	"#FFFAA0",	"#F8DE7E",	"#FDDA0D",	"#E1C16E"))+
  labs(
    x = NULL,
    y = "Relative Abundance of Fungal Phyla",
    fill = NULL
  ) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) 
```
###Protists
```{r protist relative abundances}
# Summarize protists at level 3 (various levels for inconsistent taxonomy) #
input_prot_ra_fp = filter_data(input_prot_ra, filter_cat = "sampleID", filter_vals = "F.10.A_F")
sumtax_prot <- summarize_taxonomy(input_prot_ra_fp, level = 3, report_higher_tax = F)

# clean up names... 
prot_taxonomy <- as.data.frame(rownames(sumtax_prot))
prot_taxonomy$label <- prot_taxonomy$`rownames(sumtax_prot)`

prot_taxonomy$label = recode(prot_taxonomy$label,
                               "Nucleariidae and Fonticula group" = "Rotosphaerida",
                               "Opisthokonta incertae sedis" = "Opisthokonta_incertae_sedis",
                               "Oxnerella <Centroheliozoa>" = "Centroplasthelida",
                               "PX clade" = "PX_clade",
                               "unclassified Heterolobosea" = "unclassified_Heterolobosea",
                               "unclassified Rhizaria" = "unclassified_Rhizaria",
                               "unclassified stramenopiles" = "unclassified_stramenopiles")
# go by labels, flip
rownames(sumtax_prot) <- prot_taxonomy$label
sumtax_prot.t <- as.data.frame(t(sumtax_prot))

#map
prot_map_model <- input_prot_ra$map_loaded

# merge family table and map
prot_wmap <- merge(sumtax_prot.t, prot_map_model, by = 0)

# convert to matrix for CLR 
prot_matrix <- as.matrix(prot_wmap[, 2:38])

# add small pseudocount
prot_clr <- clr(prot_matrix + 1e-6)

full_prot_clr <- bind_cols(
  prot_wmap[, -c(2:38)],
  as.data.frame(prot_clr)
)

# models for fungal abundance
# go through all variables using lmer and anova 
protist_models <- list(
Apicomplexa=lmer(Apicomplexa~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Apusomonadidae=lmer(Apusomonadidae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Bacillariophyta=lmer(Bacillariophyta~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Bangiophyceae=lmer(Bangiophyceae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Bicosoecida=lmer(Bicosoecida~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Cercozoa=lmer(Cercozoa~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Choanoflagellida=lmer(Choanoflagellida~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Chrysophyceae=lmer(Chrysophyceae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Ciliophora=lmer(Ciliophora~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Colpodellidae=lmer(Colpodellidae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Cryptomonadales=lmer(Cryptomonadales~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Dinophyceae=lmer(Dinophyceae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Discosea=lmer(Discosea~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Euglenoida=lmer(Euglenoida~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Eustigmatophyceae=lmer(Eustigmatophyceae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Foraminifera=lmer(Foraminifera~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Heterophryidae=lmer(Heterophryidae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Histionidae=lmer(Histionidae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Hyphochytriomycetes=lmer(Hyphochytriomycetes~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Kinetoplastida=lmer(Kinetoplastida~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Mycetozoa=lmer(Mycetozoa~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Rotosphaerida=lmer(Rotosphaerida~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Oikomonadaceae=lmer(Oikomonadaceae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Oomycetes=lmer(Oomycetes~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Opisthokonta_incertae_sedis=lmer(Opisthokonta_incertae_sedis~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Centroplasthelida=lmer(Centroplasthelida~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Phalansterium=lmer(Phalansterium~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
PX_clade=lmer(PX_clade~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Pyrenomonadales=lmer(Pyrenomonadales~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Raphidophyceae=lmer(Raphidophyceae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Schizopyrenida=lmer(Schizopyrenida~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Synurophyceae=lmer(Synurophyceae~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
Tubulinea=lmer(Tubulinea~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
unclassified_Heterolobosea=lmer(unclassified_Heterolobosea~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
unclassified_Rhizaria=lmer(unclassified_Rhizaria~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr),
unclassified_stramenopiles=lmer(unclassified_stramenopiles~ habitat*gopher_tortoise_activity + (1|site), data =  full_prot_clr)
)

protist_results <- do.call(rbind, lapply(names(protist_models), function(resp) {

  aov_tab <- anova(protist_models[[resp]])

  data.frame(
    response = resp,
    term     = rownames(aov_tab),
    NumDF    = aov_tab$NumDF,
    DenDF    = aov_tab$DenDF,
    F_value  = aov_tab$`F value`,
    p_value  = aov_tab$`Pr(>F)`
  )
}))

protist_results$p_fdr <- NA

for (t in unique(protist_results$term)) {
  idx <- protist_results$term == t
  protist_results$p_fdr[idx] <- p.adjust(protist_results$p_value[idx], method = "fdr")
}

protist_results2 <- protist_results %>%
  mutate(term = recode(term,
    "gopher_tortoise_activity" = "Activity",
    "habitat" = "Vegetation Type",
    "habitat:gopher_tortoise_activity" = "Interaction"
  ))

write.csv(protist_results2, "raw_protist_model_output.csv")

protist_results2 <- protist_results2 %>%
  mutate(F_p = sprintf("F = %.2f, p = %.3f", F_value, p_fdr))

protist_table <- protist_results2 %>%
  select(response, term, F_p) %>%
  pivot_wider(
    names_from = term,
    values_from = F_p
  )

protist_table <- protist_table %>%
  rename(group = response)

write.csv(protist_table, "protist_model_output_clr.csv")
```

```{r protist relative abundance figure}
# filter out F.10.A_F - it causes issues of NAs
input_prot_ra_fp = filter_data(input_prot_ra, filter_cat = "sampleID", filter_vals = "F.10.A_F")

prot_map = input_prot_ra_fp$map_loaded

## BY Treatment
table_prot_level3_fp_sort = summarize_taxonomy(input_prot_ra_fp, level = 3,
                                               report_higher_tax = F)



total_prot_sort = table_prot_level3_fp_sort %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  left_join(prot_map) %>%
  pivot_longer(cols=c(2:38), names_to = "Phylum", values_to = 'abund') %>%
  as.data.frame() 

# make mean RA by trt
trt_prot = total_prot_sort %>%
  group_by(treatment, Phylum) %>% 
  summarise(mean_RA = mean(abund))

trt_prot$Phylum = factor(trt_prot$Phylum, 
                         levels = c("Colpodellidae",
                                    "Oikomonadaceae",
                                    "Euglenoida",
                                    "Apusomonadidae",
                                    "named eukaryotic clones",
                                    "Nucleariidae and Fonticula group",
                                    "Heterophryidae",
                                    "Histionidae",
                                    "Oxnerella <Centroheliozoa>",
                                    "Discosea",
                                    "Pyrenomonadales",
                                    "Phalansterium",
                                    "PX clade",
                                    "unclassified Rhizaria",
                                    "unclassified Heterolobosea",
                                    "Mycetozoa",
                                    "Kinetoplastida",
                                    "Choanoflagellida",
                                    "Tubulinea",
                                    "Oomycetes",
                                    "Schizopyrenida",
                                    "Opisthokonta incertae sedis",
                                    "Dinophyceae",
                                    "Foraminifera",
                                    "Bacillariophyta",
                                    "Raphidophyceae",
                                    "Bangiophyceae",
                                    "Eustigmatophyceae",
                                    "Cryptomonadales",
                                    "Bicosoecida",
                                    "Apicomplexa",
                                    "Synurophyceae",
                                    "Hyphochytriomycetes",
                                    "Ciliophora",
                                    "unclassified stramenopiles",
                                    "Chrysophyceae",
                                    "Cercozoa"))

trt_prot$treatment = factor(trt_prot$treatment, 
                      levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))

ggplot(trt_prot, aes(x=treatment, y=mean_RA)) + 
  geom_bar(aes(fill = Phylum), position = "stack", stat="identity") +
  scale_fill_manual(values = c( "#004d00", "#80e680",  "#006400",  "#008f11",  "#b3ffb3",  "#00a000",  "#007a33",  "#00b300",  "#33cc33",  "#99f099",  "#ccffcc",  "#454B1B",  "#99e699",
                                "#E0B0FF",	"#7F00FF",
                               "#00004d",	"#00FFFF",	"#000066",	"#0000FF",	"#0096FF",	"#6699ff",  "#0033ff",  "#3366ff",  "#1434A4",	
                               "#4d0000", "#b30000",  "#660000",  "#ff3333",  "#cc0000",  "#ff8080",  "#990000", 
                               "#BF40BF",	"#AA336A",  "#FF00FF",	"#FF69B4",	"#FFB6C1",  "#E30B5C"
                               ))+
  labs(
    x = NULL,
    y = "Relative Abundance of Protist Taxa",
    fill = NULL
  ) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) 
```
###Nematodes
```{r nematode relative abundances}
# Summarize at the family level #
sumtax_nem_family <- summarize_taxonomy(input_nem_ra, level = 4, report_higher_tax = F)

# clean up names... create separate family level taxonomy
nem_family_taxonomy <- as.data.frame(rownames(sumtax_nem_family))
nem_family_taxonomy$label <- nem_family_taxonomy$`rownames(sumtax_nem_family)`

# go by labels, flip
rownames(sumtax_nem_family) <- nem_family_taxonomy$label
sumtax_nem_family.t <- as.data.frame(t(sumtax_nem_family))

#map
nem_map_model <- input_nem_ra$map_loaded

# merge family table and map
nem_family_wmap <- merge(sumtax_nem_family.t, nem_map_model, by = 0)

# convert to matrix for CLR 
nem_matrix <- as.matrix(nem_family_wmap[, 2:62])

# add small pseudocount
nem_clr <- clr(nem_matrix + 1e-6)

full_nem_clr <- bind_cols(
  nem_family_wmap[, -c(2:62)],
  as.data.frame(nem_clr)
)

# models for fungal abundance
# go through all variables using lmer and anova 
nematode_models <- list(
Cyatholaimidae=lmer(Cyatholaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Mermithidae=lmer(Mermithidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Trichostrongylidae=lmer(Trichostrongylidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Alaimidae=lmer(Alaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Bastianiidae=lmer(Bastianiidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Bunonematidae=lmer(Bunonematidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Cephalobidae=lmer(Cephalobidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Chronogastridae=lmer(Chronogastridae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Comesomatidae=lmer(Comesomatidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Cryptonchidae=lmer(Cryptonchidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Diplogasteridae=lmer(Diplogasteridae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Diplopeltidae=lmer(Diplopeltidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Drilonematidae=lmer(Drilonematidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Haliplectidae=lmer(Haliplectidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Isolaimiidae=lmer(Isolaimiidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Microlaimidae=lmer(Microlaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Monhysteridae=lmer(Monhysteridae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Neodiplogasteridae=lmer(Neodiplogasteridae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Odontolaimidae=lmer(Odontolaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Plectidae=lmer(Plectidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Prismatolaimidae=lmer(Prismatolaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Rhabditidae=lmer(Rhabditidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Rhabdolaimidae=lmer(Rhabdolaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Allantonematidae=lmer(Allantonematidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Steinernematidae=lmer(Steinernematidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Aphelenchidae=lmer(Aphelenchidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Aphelenchoididae=lmer(Aphelenchoididae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Diphtherophoridae=lmer(Diphtherophoridae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Leptonchidae=lmer(Leptonchidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Tylencholaimellidae=lmer(Tylencholaimellidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Tylencholaimidae=lmer(Tylencholaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Aporcelaimidae=lmer(Aporcelaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Chrysonematidae=lmer(Chrysonematidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Dorylaimidae=lmer(Dorylaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Mydonomidae=lmer(Mydonomidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Nordiidae=lmer(Nordiidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Qudsianematidae=lmer(Qudsianematidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Thornenematidae=lmer(Thornenematidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Anguinidae=lmer(Anguinidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Belondiridae=lmer(Belondiridae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Belonolaimidae=lmer(Belonolaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Criconematidae=lmer(Criconematidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Hemicycliophoridae=lmer(Hemicycliophoridae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Hoplolaimidae=lmer(Hoplolaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Longidoridae=lmer(Longidoridae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Meloidogynidae=lmer(Meloidogynidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Merliniidae=lmer(Merliniidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Pratylenchidae=lmer(Pratylenchidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Telotylenchidae=lmer(Telotylenchidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Trichodoridae=lmer(Trichodoridae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Tylenchulidae=lmer(Tylenchulidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Actinolaimidae=lmer(Actinolaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Discolaimidae=lmer(Discolaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Anatonchidae=lmer(Anatonchidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Ironidae=lmer(Ironidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Mononchidae=lmer(Mononchidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Mylonchulidae=lmer(Mylonchulidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Nygolaimidae=lmer(Nygolaimidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Tobrilidae=lmer(Tobrilidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Trischistomatidae=lmer(Trischistomatidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr),
Tylenchidae=lmer(Tylenchidae~ habitat*gopher_tortoise_activity + (1|site), data = full_nem_clr)
)

nematode_results <- do.call(rbind, lapply(names(nematode_models), function(resp) {

  aov_tab <- anova(nematode_models[[resp]])

  data.frame(
    response = resp,
    term     = rownames(aov_tab),
    NumDF    = aov_tab$NumDF,
    DenDF    = aov_tab$DenDF,
    F_value  = aov_tab$`F value`,
    p_value  = aov_tab$`Pr(>F)`
  )
}))

nematode_results$p_fdr <- NA

for (t in unique(nematode_results$term)) {
  idx <- nematode_results$term == t
  nematode_results$p_fdr[idx] <- p.adjust(nematode_results$p_value[idx], method = "fdr")
}

nematode_results2 <- nematode_results %>%
  mutate(term = recode(term,
    "gopher_tortoise_activity" = "Activity",
    "habitat" = "Vegetation Type",
    "habitat:gopher_tortoise_activity" = "Interaction"
  ))

write.csv(nematode_results2, "raw_nematode_model_output.csv")

nematode_results2 <- nematode_results2 %>%
  mutate(F_p = sprintf("F = %.2f, p = %.3f", F_value, p_fdr))

nematode_table <- nematode_results2 %>%
  select(response, term, F_p) %>%
  pivot_wider(
    names_from = term,
    values_from = F_p
  )

nematode_table <- nematode_table %>%
  rename(group = response)

write.csv(nematode_table, "nematode_model_output_clr.csv")
```

```{r nematode relative abundance figure}
nem_map = input_nem_ra$map_loaded


## BY Treatment
table_nem_level4_fp_sort = summarize_taxonomy(input_nem_ra, level = 4,
                                         report_higher_tax = T)



total_nem_sort = table_nem_level4_fp_sort %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sampleID") %>%
  left_join(nem_map) %>%
  pivot_longer(cols=c(2:62), names_to = "Family", values_to = 'abund') %>%
  separate(Family, into = c("blank", "xx", "trophic", "family"), sep = "; ") %>%
  as.data.frame() 

# make mean RA by treatment
trt_nem = total_nem_sort %>%
  group_by(treatment, family) %>% 
  summarise(mean_RA = mean(abund))

trt_nem$family = factor(trt_nem$family, 
                          levels = c("Mermithidae", # AP
                                     "Trichostrongylidae", # AP 
                                     "Cyatholaimidae", # AF 
                                     "Allantonematidae", # EMP
                                     "Tylenchidae", # RA
                                     "Tobrilidae", ## end of PR
                                     "Ironidae",
                                     "Discolaimidae",
                                     "Actinolaimidae",
                                     "Nygolaimidae",
                                     "Mylonchulidae",
                                     "Trischistomatidae",
                                     "Anatonchidae",
                                     "Mononchidae",
                                     "Chrysonematidae", ## end of OM
                                     "Nordiidae",
                                     "Mydonomidae",
                                     "Thornenematidae",
                                     "Aporcelaimidae",
                                     "Qudsianematidae",
                                     "Dorylaimidae",
                                     "Trichodoridae", ## end of PP
                                     "Hoplolaimidae",
                                     "Hemicycliophoridae",
                                     "Meloidogynidae",
                                     "Belondiridae",
                                     "Pratylenchidae",
                                     "Tylenchulidae",
                                     "Merliniidae",
                                     "Belonolaimidae",
                                     "Longidoridae",
                                     "Criconematidae",
                                     "Telotylenchidae",
                                     "Anguinidae",
                                     "Tylencholaimellidae", ## end of FF
                                     "Aphelenchidae",
                                     "Tylencholaimidae",
                                     "Leptonchidae",
                                     "Aphelenchoididae",
                                     "Diphtherophoridae",
                                     "Isolaimiidae", ## end of BF
                                     "Diplogasteridae", 
                                     "Bunonematidae",
                                     "Diplopeltidae",
                                     "Haliplectidae",
                                     "Comesomatidae",
                                     "Chronogastridae",
                                     "Drilonematidae",
                                     "Steinernematidae",
                                     "Rhabditidae",
                                     "Alaimidae",
                                     "Microlaimidae",
                                     "Odontolaimidae",
                                     "Bastianiidae",
                                     "Rhabdolaimidae",
                                     "Neodiplogasteridae",
                                     "Cryptonchidae",
                                     "Monhysteridae",
                                     "Prismatolaimidae",
                                     "Plectidae",
                                     "Cephalobidae"))


trt_nem$treatment = factor(trt_nem$treatment, 
                            levels = c("IntactBurrowing", "IntactForaging", "DegradedBurrowing", "DegradedForaging"))

ggplot(trt_nem, aes(x=treatment, y=mean_RA)) + 
  geom_bar(aes(fill = family), position = "stack", stat="identity") +
  scale_fill_manual(values = c("#E0B0FF",	"#7F00FF",		
                               "#40E0D0",
                               "grey", 
                               "black",
                               "#00004d",	"#00FFFF",	"#000066",	"#0000FF",	"#0096FF",	"#6699ff",  "#0033ff",  "#3366ff",  "#1434A4",	
                               "#4d0000", "#b30000",  "#660000",  "#ff3333",  "#cc0000",  "#ff8080",  "#990000", 
                               "#004d00", "#80e680",  "#006400",  "#008f11",  "#b3ffb3",  "#00a000",  "#007a33",  "#00b300",  "#33cc33",  "#99f099",  "#ccffcc",  "#454B1B",  "#99e699", 
                               "#BF40BF",	"#AA336A",  "#FF00FF",	"#FF69B4",	"#FFB6C1",  "#E30B5C",
                               "#FFBF00",	"#FBCEB1",	"#CD7F32",	"#FFC000",	"#FFEA00",	"#F28C28",	"#E97451",	"#FAD5A5",	"#DAA06D",	"#F4BB44",	"#FFAA33",	"#B87333",	"#FF7518",	"#CC5500",	"#FFFF8F",	"#F0E68C",	"#FADA5E",	"#FFFAA0",	"#F8DE7E",	"#FDDA0D",	"#E1C16E"))+
  labs(
    x = NULL,
    y = "Relative Abundance of Nematode Families",
    fill = NULL
  ) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  guides(fill = guide_legend(nrow = 61))
```
##Last analyses is looking at the distance based redundancy analyses for each major group 
###Deciding whether to use diversity or richness as metrics in the analyses using correlations
```{r correlations of all diversity metrics}
biotic_cormap<-input_nem_ra$map_loaded[,43:50] 
M=cor(biotic_cormap,use="pairwise.complete.obs")
corrplot(M, method = 'number', type = 'upper',number.cex = 0.75,tl.cex = 0.75)
```
##We will use richness going forwarsd
##These dbRDAs are split between the different habitat types
###Start with bacteria 
```{r bacteria dbRDA}
## bacterial dbRDA for intact
input_bact_ra_i = filter_data(input_bact_ra, filter_cat = "habitat", keep_vals = "Intact")
bray = vegdist(t(input_bact_ra_i$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+fung_rich+prot_rich+mean_plant_rich,data=input_bact_ra_i$map_loaded, na.action = na.fail)
# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_bact_ra_i$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_bact_ra_i$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(21, 21)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("IntactBurrowing", "IntactForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)

## bacterial dbRDA for degraded
input_bact_ra_d = filter_data(input_bact_ra, filter_cat = "habitat", keep_vals = "Degraded")
bray = vegdist(t(input_bact_ra_d$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+fung_rich+prot_rich+mean_plant_rich,data=input_bact_ra_d$map_loaded, na.action = na.fail)
# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_bact_ra_d$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_bact_ra_d$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(25, 25)  # Different point shapes for different treatments
colvec <- c("khaki3","green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("DegradedBurrowing", "DegradedForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)
```
###Then fungi
```{r fungi dbRDA}
# intact
input_fung_ra_i = filter_data(input_fung_ra, filter_cat = "habitat", keep_vals = "Intact")
bray = vegdist(t(input_fung_ra_i$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+prot_rich+mean_plant_rich,data=input_fung_ra_i$map_loaded, na.action = na.fail)
# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_fung_ra_i$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_fung_ra_i$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(21, 21)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("IntactBurrowing", "IntactForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)

## fungal dbRDA for degraded
input_fung_ra_d = filter_data(input_fung_ra, filter_cat = "habitat", keep_vals = "Degraded")
bray = vegdist(t(input_fung_ra_d$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+prot_rich+mean_plant_rich,data=input_fung_ra_d$map_loaded, na.action = na.fail)
# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_fung_ra_d$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)


map = input_fung_ra_d$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(25, 25)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("DegradedBurrowing", "DegradedForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)
```
###Then protists
```{r protist dbRDA}
# intact
input_prot_ra_i = filter_data(input_prot_ra, filter_cat = "habitat", keep_vals = "Intact")
bray = vegdist(t(input_prot_ra_i$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+mean_plant_rich,data=input_prot_ra_i$map_loaded, na.action = na.fail)
# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_prot_ra_i$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_prot_ra_i$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(21, 21)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("IntactBurrowing", "IntactForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)

## protist dbRDA for degraded
input_prot_ra_d_fp = filter_data(input_prot_ra, filter_cat = "sampleID", filter_vals = "F.10.A_F")
input_prot_ra_d = filter_data(input_prot_ra_d_fp, filter_cat = "habitat", keep_vals = "Degraded")
bray = vegdist(t(input_prot_ra_d$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+mean_plant_rich,data=input_prot_ra_d$map_loaded, na.action = na.fail)
# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_prot_ra_d$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)


map = input_prot_ra_d$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(25, 25)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("DegradedBurrowing", "DegradedForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)
```
###Nematode dbRDAs split by trophic group
```{r bacterial-feeding}
## using only intact samples
input_BF = filter_taxa_from_input(input_nem_ra, taxa_to_keep = "BF")
input_BF_filt = filter_data(input_BF, filter_cat = 'sampleID',
                            filter_vals = c("F.7.A_F", "S.3.A_F"))
input_BF_filt_i = filter_data(input_BF_filt, filter_cat = "habitat",
                              keep_vals = "Intact")
bray = vegdist(t(input_BF_filt_i$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+prot_rich+mean_plant_rich,data=input_BF_filt_i$map_loaded, na.action = na.omit)

# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_BF_filt_i$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_BF_filt_i$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(21, 21)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("IntactBurrowing","IntactForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)

## using only degraded samples
input_BF_filt_d = filter_data(input_BF_filt, filter_cat = "habitat",
                              keep_vals = "Degraded")
bray = vegdist(t(input_BF_filt_d$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+prot_rich+mean_plant_rich,data=input_BF_filt_d$map_loaded, na.action = na.omit)

# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_BF_filt_d$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_BF_filt_d$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(25, 25)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("DegradedBurrowing","DegradedForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)
```

```{r fungal-feeding}
## using only intact samples
input_FF = filter_taxa_from_input(input_nem_ra, taxa_to_keep = "FF")
input_FF_filt = filter_data(input_FF, filter_cat = 'sampleID',
                            filter_vals = c("S.7.A_F", "S.8.A_F", "F.5.A_F",
                                            "S.6.O_F", "F.7.A_F", "S.3.A_F"))
input_FF_filt_i = filter_data(input_FF_filt, filter_cat = "habitat",
                              keep_vals = "Intact")
bray = vegdist(t(input_FF_filt_i$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+prot_rich+mean_plant_rich,data=input_FF_filt_i$map_loaded, na.action = na.omit)

# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_FF_filt_i$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_FF_filt_i$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(21, 21)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("IntactBurrowing","IntactForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)

## using only degraded samples
input_FF_filt_d = filter_data(input_FF_filt, filter_cat = "habitat",
                              keep_vals = "Degraded")
bray = vegdist(t(input_FF_filt_d$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+prot_rich+mean_plant_rich,data=input_FF_filt_d$map_loaded, na.action = na.omit)

# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_FF_filt_d$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_FF_filt_d$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(25, 25)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("DegradedBurrowing","DegradedForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)
```

```{r omnivorous}
## using only intact samples
input_OM = filter_taxa_from_input(input_nem_ra, taxa_to_keep = "OM")
input_OM_filt = filter_data(input_OM, filter_cat = 'sampleID',
                            filter_vals = c("S.9.O_F", "F.8.A_F", "F.5.A_F",
                                            "S.3.A_F", "F.7.A_F", "S.6.O_F"))
input_OM_filt_i = filter_data(input_OM_filt, filter_cat = "habitat",
                              keep_vals = "Intact")
bray = vegdist(t(input_OM_filt_i$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+prot_rich+mean_plant_rich,data=input_OM_filt_i$map_loaded, na.action = na.omit)

# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_OM_filt_i$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_OM_filt_i$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(21, 21)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("IntactBurrowing","IntactForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)

## using only degraded samples
input_OM_filt_d = filter_data(input_OM_filt, filter_cat = "habitat",
                              keep_vals = "Degraded")
bray = vegdist(t(input_OM_filt_d$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+prot_rich+mean_plant_rich,data=input_OM_filt_d$map_loaded, na.action = na.omit)

# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_OM_filt_d$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_OM_filt_d$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(25, 25)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("DegradedBurrowing","DegradedForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)
```

```{r plant-parasitic}
## using only intact samples
input_PP = filter_taxa_from_input(input_nem_ra, taxa_to_keep = "PP")
input_PP_filt = filter_data(input_PP, filter_cat = 'sampleID',
                            filter_vals = c("S.6.O_F"))
input_PP_filt_i = filter_data(input_PP_filt, filter_cat = "habitat",
                              keep_vals = "Intact")
bray = vegdist(t(input_PP_filt_i$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+prot_rich+mean_plant_rich,data=input_PP_filt_i$map_loaded, na.action = na.omit)


# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_PP_filt_i$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_PP_filt_i$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(21, 21)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("IntactBurrowing","IntactForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)

## using only degraded samples
input_PP_filt_d = filter_data(input_PP_filt, filter_cat = "habitat",
                              keep_vals = "Degraded")
bray = vegdist(t(input_PP_filt_d$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+prot_rich+mean_plant_rich,data=input_PP_filt_d$map_loaded, na.action = na.omit)

# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_PP_filt_d$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_PP_filt_d$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(25, 25)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("DegradedBurrowing","DegradedForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)
```

```{r root-associated}
## using only intact samples
input_RA = filter_taxa_from_input(input_nem_ra, taxa_to_keep = "RA")
input_RA_filt = filter_data(input_RA, filter_cat = 'sampleID',
                            filter_vals = c("F.1.A_F", "F.3.A_F", "F.3.A_F",
                                            "F.5.A_F", "F.6.A_F", "F.9.A_F",
                                            "F.13.A_F", "F.1.O_F", "F.3.O_F",
                                            "F.4.O_F", "F.5.O_F", "F.6.O_F",
                                            "F.8.O_F", "F.9.O_F", "S.2.A_F",
                                            "S.4.A_F", "S.5.A_F", "S.8.A_F",
                                            "S.12.A_F", "S.9.O_F", "S.3.A_F",
                                            "S.6.O_F", "F.7.A_F"))
input_RA_filt_i = filter_data(input_RA_filt, filter_cat = "habitat",
                              keep_vals = "Intact")
bray = vegdist(t(input_RA_filt_i$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+prot_rich+mean_plant_rich,data=input_RA_filt_i$map_loaded, na.action = na.omit)

# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_RA_filt_i$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_RA_filt_i$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(21, 21)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("IntactBurrowing","IntactForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)

## using only degraded samples
input_RA_filt_d = filter_data(input_RA_filt, filter_cat = "habitat",
                              keep_vals = "Degraded")

bray = vegdist(t(input_RA_filt_d$data_loaded))

# build model
dbrda.m1<-dbrda(bray ~ gopher_tortoise_activity+Soil.pH+Nitrate+Ammonium+Calcium+Organic.Matter+Total.Organic.C+Gravimetric.H2O+Total.Phosphorus+bact_rich+fung_rich+prot_rich+mean_plant_rich,data=input_RA_filt_d$map_loaded, na.action = na.omit)

# test significance of model terms
anova.cca(dbrda.m1, by='term')

## check for homogeneity
disp_dbrda <- betadisper(bray, input_RA_filt_d$map_loaded$gopher_tortoise_activity)
anova(disp_dbrda)

map = input_RA_filt_d$map_loaded

# Define point shapes and colors based on treatment group
pchc <- c(25, 25)  # Different point shapes for different treatments
colvec <- c("khaki3", "green4")  # Colors corresponding to treatments
scl <- 2  # Scaling factor for the dbRDA plot

# Create an empty plot to set up the axes limits
plot(dbrda.m1, type = "n", scaling = scl, xlim = c(-4, 4), ylim = c(-3, 3))


# Add text labels for the biplot (representing the explanatory variables)
text(dbrda.m1, col = "gray", display = "bp")

# Add a legend to the plot with point shapes and colors representing treatments
map$treatment<-as.character(map$treatment)  
map$treatment<-factor(map$treatment, levels=c("DegradedBurrowing","DegradedForaging"))
with(map, levels(map$treatment))

map$treatment <- as.factor(map$treatment)

with(map, 
     legend(-4.2, 2, 
            legend = levels(treatment),  # Treatment group names
            bty = "n",             # No box around legend
            col = colvec,          # Colors for treatments
            pch = pchc,            # Point shapes for treatments
            pt.bg = colvec         # Background colors for points in legend
     ))

# Add points for the sites using the treatment (treatment) groups from 'map'
with(map, 
     points(dbrda.m1, 
            display = "sites", 
            col = colvec[treatment],  # Color based on treatment variable
            scaling = scl, 
            pch = pchc[treatment],    # Shape based on treatment variable
            bg = colvec[treatment])   # Background color for points
)
```
